Mirror, mirror on the wall. Who's the best match of them all?

### 2023-11-22 PNAS/index.Rnw
references, images

\firstpage[3]{12}
% Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting. In this example, there are a total of 11 paragraphs on the first page, counting the first level heading as a paragraph. The value {12} represents the number of the paragraph starting the second page. If a paragraph runs over onto the second page, include a bracket with the paragraph line number starting the second page, followed by the paragraph number in curly brackets, e.g. "\firstpage[4]{11}".

@@

This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehlerFingerprintErrorRate2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
%
\hh{In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.}
%
\hh{In order to avoid embarassing false accusation and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.}
%
\hh{Database searches are an example where the number of searches is explicitly available as the number of objects meeting the search criteria. }

@@

studies <- tibble(single_rmp = c( 0.004545,0.00704, 0.0724), study = c("Best (2022)","Bajic (2020)",  "Mattijssen (2021)"), y = c(-0.025, .15, 0.6), hjust = c(0, 0, 1))
ggplot(res, aes(x = single_rmp, y = overall, group = factor(N), color = factor(N))) +
  geom_line() +
  coord_cartesian(xlim = c(0, 0.07), ylim = c(0, 1)) +
  ylab("Probability of a coincidental match (N comparisons)") +
  xlab("Probability of a false positive (1 comparison)") +
  scale_y_continuous(breaks = seq(0, 1, .1), minor_breaks = seq(0, 1, .05)) +
  ggtitle("Wire Comparison: Controlling Overall False Positive Rate") +
  geom_vline(xintercept = 0.02, linetype = "dashed", color = "grey40") +
  geom_vline(xintercept = studies$single_rmp, linetype = "51", color = "grey50") +
  geom_label(data = studies[1,], aes(x = single_rmp, y = y, label = study), inherit.aes = F, hjust = studies$hjust[1], vjust = 0) +
  geom_label(data = studies[2,], aes(x = single_rmp, y = y, label = study), inherit.aes = F, hjust = studies$hjust[2], vjust = 0) +
  geom_label(data = studies[3,], aes(x = single_rmp, y = y, label = study), inherit.aes = F, hjust = studies$hjust[3], vjust = 0) +
  annotate(geom = "label", x = 0.02, y = 0.1, label = "Pooled error rate \nused in this paper", color = "grey40", hjust=0) +
  scale_color_discrete("# comparisons") +
  theme_bw() +
  theme(legend.position = c(1, 0), legend.justification = c(1, 0), legend.background = element_rect(color = "black"))
@

@@

ggplot(data = res2, aes(x = N, y = overall, color = factor(r), label = rlab)) +
  geom_textline(parse = T, gap = F, straight = F, padding = unit(1, "mm"), vjust = 0, linewidth = 0.75) +
  geom_line(data = error_res, linetype = 2, linewidth = 0.75) +
  xlab("Number of comparisons") +
  scale_x_log10() +
  geom_hline(yintercept = breakpt, linewidth = 1) +
  geom_label(data = res2sum, aes(x = N, y = breakpt, label = paste0("N == ", round(N))), parse = T, vjust = 1, hjust = 1) +
  ylab("Probability of no coincidental matches (N comparisons)") +
  ggtitle("# Comparisons and Overall Accuracy") +
  scale_color_discrete("FP Rate for single comparison", guide = 'none') +
  theme_bw() +
  coord_cartesian(ylim = c(0.75, 1.0)) +
  theme(axis.text.y = element_text(hjust = 0.5, vjust = 0.5, angle = 90))

>>>>

\firstpage[11]{6}
% Use \firstpage to indicate which paragraph and line will start the second page and subsequent formatting. In this example, there are a total of 11 paragraphs on the first page, counting the first level heading as a paragraph. The value {12} represents the number of the paragraph starting the second page. If a paragraph runs over onto the second page, include a bracket with the paragraph line number starting the second page, followed by the paragraph number in curly brackets, e.g. "\firstpage[4]{11}".

@@

This issue has been raised in the past with regard to DNA \citep{thompson2003} and fingerprints \citep{koehler2021}: one of the issues \citep{fine2006} leading to the wrongful accusation of Brandon Mayfield in the 2004 Madrid bombing case was that the large size of the IAFIS database used to locate similar prints made it possible to locate `unusually' similar non-matches.
%
\hh{In this paper, we discuss that very close non-matches are far from an unusual phenomenon  but a direct consequence of a large number of comparisons, and simply reflect the power of the search algorithms.}
%
\hh{In order to avoid embarassing false accusation and a further erosion of the public's trust in science, we need to address the problem of multiple comparisons by controlling their effect on false discovery rates.}
%
\hh{Database searches are an example where the number of searches is explicitly available as the number of objects meeting the search criteria. }

@@

studies <- tibble(single_rmp = c( 0.004545,0.00704, 0.0724), study = c("Best (2022)","Bajic (2020)",  "Mattijssen (2021)"), y = c(-0.025, .15, 0.6), hjust = c(0, 0, 1))
ggplot(res, aes(x = single_rmp, y = overall, group = factor(N), color = factor(N))) +
  coord_cartesian(xlim = c(0, 0.07), ylim = c(0, 1)) +
  ylab("Probability of a false positive\n(N comparisons)") +
  xlab("Probability of a false positive\n(1 comparison)") +
  ggtitle("Wire Comparison: Controlling Overall False Positive Rate") +
  geom_vline(xintercept = 0.02, linetype = "dashed", color = "grey40") +
  geom_vline(xintercept = studies$single_rmp, linetype = "51", color = "grey50") +
  geom_label(data = studies[1,], aes(x = single_rmp, y = y, label = study), inherit.aes = F, hjust = studies$hjust[1], vjust = 0) +
  geom_label(data = studies[2,], aes(x = single_rmp, y = y, label = study), inherit.aes = F, hjust = studies$hjust[2], vjust = 0) +
  geom_label(data = studies[3,], aes(x = single_rmp, y = y, label = study), inherit.aes = F, hjust = studies$hjust[3], vjust = 0) +
  annotate(geom = "label", x = 0.02, y = 0.1, label = "Pooled error rate \nused in this paper", color = "grey40", hjust=0) +
  scale_color_discrete("# comparisons  ") +
  geom_line(linewidth=1) +
  theme_bw() +
  theme(legend.position = c(1, 0),
        legend.justification = c(1, 0),
        legend.background = element_rect(color = "grey40", fill=alpha("white", 0.1))) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), minor_breaks = seq(0, 1, .05)) +
  scale_x_continuous(breaks = c(0, 0.02, 0.04, 0.06, studies$single_rmp),
                     labels=c(0, 0.02, 0.04, 0.06, "0.004545", "\n0.00704","0.0724"))
#  geom_hline(yintercept = 0.1, colour="black")
@

@@

ggplot(data = res2, aes(x = N, y = overall, color = factor(r), label = rlab)) +
  geom_textline(parse = T, gap = F, straight = F,
                padding = unit(1, "mm"), vjust = 0, linewidth = 1) +
  xlab("Number of comparisons") +
  scale_x_log10(
    breaks = c(1, 1000, 10000, res2sum$N),
    minor_breaks = c((1:9)[-2],10*(3:9), 100*3:9, 1000*1:9)
  ) +
  geom_hline(yintercept = breakpt, linewidth = 1) +
#  geom_label(
#    data = res2sum,
#    aes(x = N, y = breakpt, label = paste0("N == ", round(N))),
#    parse = T, vjust = 1, hjust = 1) +
  geom_segment(data = res2sum,
     aes(x = N, xend=N, y = -Inf, yend=0.9), colour="grey30", linewidth=0.5) +
  geom_line(data = error_res, linetype = 2, linewidth = 0.75) +
  ylab("Probability false positive\n(N comparisons)") +
  ggtitle("# Comparisons and Overall Accuracy") +
  scale_color_discrete("FP Rate for single comparison", guide = 'none') +
  theme_bw() +
  coord_cartesian(ylim = c(0.775, 1.0)) +
  theme(axis.text.y = element_text(hjust = 0.5, vjust = 0.5, angle = 90))
